### 2021-11-07 paper/output/document/whykant.tex
cleaned much of naive

%
\begin{isabellebody}%
\setisabellecontext{whykant}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Why Kantian Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this thesis, I automate Kantian ethics. In 2006, Powers posited that deontological theories are 
attractive candidates for automation because rules are generally computationally tractable \cite[1]{powers}. 
Intuitively, algorithms are rules or procedures for problem solving and Kantian ethics offer one such 
procedure for the problem of making ethical judgements. I will make this intuition precise by
arguing that Kantian ethics is natural to formalize because evaluating a maxim requires little additional
data about the world and a maxim is relatively easy to represent to a computer. All 
ethical traditions have debates that an automated ethical system will need to take a stance on, but these debates
are less frequent and less controversial for Kantian ethics than for consequentialism and 
virtue ethics.

I do not aim to show that Kantian ethics is the only tractable theory to automate and
to present a comprehensive overview of all consequentialist or virtue ethical theories. Instead, I 
present a sample of some approaches in each tradition and argue that Kantian ethics is more straightforward 
to formalize than these approaches. Future work could and should address the challenges I outline in 
this section. The more ethical theories that computational tools can handle, the more valuable 
computational philosophy becomes both for philosophers and for AI agents. Insofar as my project serves 
as an early proof-of-concept for computational ethics, I choose to automate an ethical theory that 
poses fewer challenges than others.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Kantian Ethics%
}
\isamarkuptrue%
%
\isamarkupsubsubsection{Crash Course on Kantian Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kant's theory is centered on practical reason, which is the kind of reason that we 
use to decide what to do. In \emph{The Groundwork of the Metaphysics of Morals}, Kant's most influential 
text on ethics, he explains that rational beings are unique because we can act ``in accordance with 
the representations of laws" \cite[4:412]{groundwork}. In contrast, a ball thrown into the air acts 
according to the laws of physics themselves. It cannot ask itself, ``Should I fall back to the ground?" 
It simply falls. A rational being on the other hand, can ask themselves ``Should I act on this reason?" 
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on" \cite[100]{sources}. 
Rational beings are set apart by this reflective capacity. A rational being's behavior is purposive and 
their actions are guided by practical reason. They have reasons for acting, even when these reasons may be 
opaque to them. This operation of practical reason is what Kant calls the will.

@@

The first formulation of the categorical imperative is the
Formula of Universal Law (FUL), which reads, ``act only according to that maxim through which you can 
at the same time will that it become a universal law" $\cite[34]{groundwork}$. This formulation
generates the universalizability test, which tests the moral value of a maxim by 
imagining a world in which it becomes a universal law and attempting to will the maxim in that world.
If there is a contradiction in willing the maxim in a world in which everyone universally wills the maxim,
the maxim is prohibited. 
Velleman presents a concise argument for the FUL. He argues that reason is universally shared among reasoners. For 
example, all reasoners have equal access to the arithmetic logic that shows that ``2+2=4" \cite[29]{velleman}. The chain of 
reasoning that makes this statement true is not specific to any person, but is universal across people. 
Therefore, if I have sufficient reason to will a maxim, so does every other rational agent. There is 
nothing special about the operation of my practical reason that other reasoners don't have access to. 
Practical reason is shared, so in adopting a maxim, I implicitly state that all reasoners
across time also have reason to adopt that maxim. Therefore, because I act on reasons, I must obey the 
FUL. Notice that this fulfills the above criterion for a categorical imperative: the FUL is derived from 
a property of practical reason itself and thus derives authority from the will's authority over itself, 
as opposed to some external authority.

The second formulation of the categorical imperative is the formula of humanity (FUH): ``So act that you use humanity, 
in your own person, as well as in the person of any other, always at the same time as an end, never merely 
as a means." $\cite[41]{groundwork}$. This formulation is often understood as requiring us to 
acknowledge and respect the dignity of every other person. The third formulation of the categorical 
imperative is the formula of autonomy (FOA), which Korsgaard summarizes in her introduction to the Groundwork 
as, ``we should so act that we may think of ourselves as legislating universal laws through our 
maxims" $\cite[28]{korsgaardintro}$. While closely related to the FUL, the FOA presents morality as the activity of 
perfectly rational agents in an ideal ``kingdom of ends," guided by what Kant calls the ``laws of freedom."

The above is not meant to serve as a full defense of Kant's ethical theory, as that is outside the scope
of this thesis. Instead, I briefly reconstruct an argument for Kant's ethical theory in the hopes 
of offering context for the implementation of the FUL I present later in the thesis. Additionally, understanding 
the structure of Kant's argument also reveals facts about his theory that make it an ideal candidate 
for formalization.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Ease of Automation%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kantian ethics is natural to formalize because the categorical imperative, particularly the FUL, 
is a formal principle of practical reason. In other words, the FUL is a property of reason related 
to the form or structure of a maxim. It has nothing to do with the circumstances of behavior 
(beyond those included in a maxim), the agent's mental state, or any other contingent facts. Instead, 
it is purely a property of the proposed principle for action. This formalism makes Kantian ethics an 
attractive candidate for formalization. While other ethical theories often rely on many facts about 
the world or the actor, Kantian ethics simply relies on the form of a given maxim. A computer evaluating 
a maxim doesn't require any knowledge about the world beyond what is contained in a maxim. A maxim 
is the only input that the computer needs to make a moral judgement. Automating 
Kantian ethics merely requires making the notion of a maxim precise and representing it to the computer. 
This distinguishes Kantian ethics from consequentialism and virtue ethics, which, as I will argue below, 
require far more knowledge about the world or the agent to reach a moral decision.

@@

I do not argue that Kantian ethics is the only theory possible to formalize, but instead that Kantian
ethics is relatively easier to formalize. Indeed, like any ethical theory, the Kantian tradition contains 
debates that philosophers disagree on. In this project, I assume stances on debates about
what exactly constitutes the form of a maxim and the correct interpretation of the Formula of 
Universal Law. Those who disagree with my stances will not trust the judgements of my system. Unlike
consequentialism or virtue ethics, these debates are considered close to settled in the Kantian literature 
\cite{ebelsduggan}. Because such debates are fewer and less controversial than those in virtue ethics
or consequentialism, Kantian ethics is relatively easier to formalize.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Consequentialism%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
A consequentialist ethical theory is, broadly speaking, any ethical theory that evaluates an action by evaluating 
its consequences. \footnote{There is long debate about what exactly makes an ethical theory consequentialist \cite{consequentialismsep}. 
For this paper, I will focus on theories that place the moral worth of an act in its the consequences.} For example, 
utilitarianism is a form of consequentialism in which the moral action 
is the action that results in the best consequences or produces the most good \cite{utilsep}. The focus
on the consequences of action distinguishes consequentialists from Kantians, who derive the moral worth
of an action from the maxim that is acted on. Some debates in the consequentialist tradition include 
which consequences of an action matter, what exactly constitutes a ``good" consequence, and how we can 
aggregate the consequences of an action over all the individuals involved.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Which Consequences Matter%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Because consequentialism evaluates the state of affairs following an action, this kind of ethical 
reasoning requires more knowledge
about the state of the world than Kantian ethics. Under a naive version of consequentialism, evaluating 
an action requires perfect knowledge of all consequences following an action. This requires that an 
automated ethical system somehow collect all of the infinite consquences following an action, a likely
impossible task. Moreover, compiling this database of consequences requires 
answering difficult questions about which consequences were actually caused by an action.\footnote
{maybe cite the debate about difficulties in determining causation?}

@@

The challenge of understanding and representing the circumstances of action is not unique to consequentialism,
but is particularly acute for consequentialism. Kantian ethicists robustly debate which circumstances 
of an action are ``morally relevant" and should be included in the formulation of the maxim.\footnote{Powers 
\cite{powers} identifies this as a challenge for automating Kantian ethics and briefly sketches 
solutions from O'Neill \cite{constofreason}, Silber \cite{silber}, and Rawls \cite{rawlsconstructivism}. } Those 
using my system will need to automatically or manually formulate maxims, a process that involves 
common-sense reasoning to determine which circumstances, act, and goal tuples the agent is adopting.
\footnote{For more on the challenge of parsing of ethical dilemmas into maxims, see Section AI Ethics}
Because Kantian ethics merely evaluates a single maxim, the surface of this debate is much smaller
than the debate about circumstances and consequences in a consequentialist system. An automated 
consequentialist system must make such judgements about the act itself, the circumstances in which 
it is performed, and the circumstances following the act. All ethical theories relativize
their judgements to the situation in which an act is performed to some extent, but consequentialism
requires far more knowledge about the world than Kantian ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Theory of the Good%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Another debate that an automated consequentialist reasoner would need to take a stance on is
the question of what qualifies as a ``good consequence," or what the theory of the good is. Hedonists associate
good with the presence of pleasure and the absence of pain. Preference utiliarians believe that good is 
the satisfaction of desires and is thus derived from individuals' preferences, as opposed to some
sensation of pleasure or pain. Other consequentialists adopt a pluralistic theory of value, under which 
many different kinds of things are good for different reasons. For example, Moore values beauty and truth 
and other pluralists value justice, love, and freedom \cite{moorepe}. Welfare utilitarians value a person's 
welfare and utilitarians of right value states of affairs in which respect for some set of
rights is maximized \cite{consequentialismsep}.

Most of the above theories of good require that a moral reasoner understand complex features about
individuals' preferences, desires, or sensations in order to evaluate a moral action, making automated
consequentialist ethics difficult. Regardless of the theory of the good, a consequentialist ethical 
reasoner needs to evaluate a state of affairs, which encompass each involved individual's pleasure, 
preferences, welfare, freedom, rights, or whatever other criteria make a state good. This requires
judgements about whether or not a state of affairs actually satisifes the relevant criteria for goodness. 
These judgements are difficult and debateable, and any consequentialist decision requires many of 
these judgements for each individual involved. As systems become more complex and involve more people and 
more acts, making these judgements quickly becomes difficult, posing a scaling challenge for a 
consequentialist ethical reasoner. Perfect knowledge of tens of thousands of people's pleasure or 
preferences or welfare or rights is impossible. Either a human being 
assigns values to states of affairs, which quickly becomes difficult to scale, or the machine does, 
which requires massive common-sense, increases room for doubting the system's judgements, and simplifies
the judgements. This is a tractable problem, but is much more difficult than the Kantian task of formulating
and evaluating a maxim.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Aggregation%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Once an automated consequentialist agent assigns a goodness measurement to each person in a state of affairs, it 
must also calculate an overall goodness measurement for the state of affairs. One approach to assigning
this value is to aggregate each person's individual goodness score into one complete score for a state. 
For example, under a simple welfare model, each person is assigned a welfare score and the total 
score for a state of affairs is the sum of the welfare scores for each involved person.
The more complex the theory of the good, the more difficult this aggregation becomes. For example, 
pluralistic theories struggle to explain how different kinds of value can be compared \cite{consequentialismsep}. 
How do we compare one unit of beauty to one unit of pleasure? Subjective theories of the good, such 
as those focused on the sensation of pleasure or an individual's preferences, present difficulties in 
comparing different people's subjective measures. Resolving this debate requires that the automated reasoner 
choose one specific aggregation algorithm, but those who disagree with this choice will not trust 
the reasoner's moral judgements. Moreover, for complex theories of the good, this aggregation algorithm
may be complex and may require a lot of data.

To solve this problem, some consequentialists reject aggregation entirely and instead prefer wholistic
evaluations of a state of affairs. While this approach no longer requires that a reasoner define an 
aggregation algorithm, the reasoner still needs to calculate a goodness measurement for a state of 
affairs. Whereas before the reasoner could restrict analysis to a single person, the algorithm must now 
evaluate an entire state wholistically. Evaluating the goodness of an entire state of affairs is more complicated
than evaluating the goodness of a single person. As consequentialists modulate between aggregation 
and wholistic evaluation, they face a tradeoff between the difficulty of aggregation and the complexity 
of goodness measurements for large states of affairs. This tradeoff also holds for an automated
consequentialist moral agent. Such an agent either needs to define an aggregation function, thus opening 
the door to those who disagree with this definition, or needs to evaluate the goodness of entire states
of affairs, which is a complex and data-intensive philosophical and technical challenge.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Prior Attempts to Formalize Consequentialism%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
None of the challenges described above are intractable or capture the full literature of 
all variations of consequentialism. I do not argue that consequentialism is 
impossible to automate. Instead, each of the challenges above requires that the developer 
``plant certain flags" and take a stance on certain philosophical debates. Such debates are present in 
any ethical theory, but consequentialism has more such points of difficulty than Kantian ethics and 
is thus more difficult to automate.

@@

\isamarkupsubsection{Virtue Ethics%
\isamarkupsubsubsection{What Is Virtue%

@@

@@

\isamarkupsubsubsection{Evaluating Moral Character%

@@

\isamarkupsubsubsection{Machine Learning and Virtue Ethics%

@@

%:%66=31%:%

@@

%:%163=111%:%
%:%175=113%:%
%:%176=114%:%
%:%177=115%:%
%:%178=116%:%
%:%179=117%:%
%:%180=118%:%
%:%181=119%:%
%:%182=120%:%
%:%183=121%:%
%:%184=122%:%
%:%185=123%:%
%:%186=124%:%
%:%187=125%:%
%:%188=126%:%
%:%189=127%:%
%:%190=128%:%
%:%191=129%:%
%:%192=130%:%
%:%193=131%:%
%:%194=132%:%
%:%195=133%:%
%:%196=134%:%
%:%197=135%:%
%:%198=136%:%
%:%199=137%:%
%:%200=138%:%
%:%201=139%:%
%:%202=140%:%
%:%203=141%:%
%:%204=142%:%
%:%205=143%:%
%:%206=144%:%
%:%207=145%:%
%:%216=148%:%
%:%228=150%:%
%:%229=151%:%
%:%230=152%:%
%:%231=153%:%
%:%232=154%:%
%:%233=155%:%
%:%234=156%:%
%:%235=157%:%
%:%236=158%:%
%:%245=160%:%
%:%257=162%:%
%:%258=163%:%
%:%259=164%:%
%:%260=165%:%
%:%261=166%:%
%:%262=167%:%
%:%263=168%:%
%:%264=169%:%
%:%265=170%:%
%:%266=171%:%
%:%267=172%:%
%:%268=173%:%
%:%269=174%:%
%:%270=175%:%
%:%271=176%:%
%:%272=177%:%
%:%273=178%:%
%:%274=179%:%
%:%275=180%:%
%:%276=181%:%
%:%277=182%:%
%:%278=183%:%
%:%279=184%:%
%:%280=185%:%
%:%281=186%:%
%:%282=187%:%
%:%283=188%:%
%:%284=189%:%
%:%285=190%:%
%:%286=191%:%
%:%287=192%:%
%:%288=193%:%
%:%289=194%:%
%:%290=195%:%
%:%291=196%:%
%:%292=197%:%
%:%293=198%:%
%:%294=199%:%
%:%295=200%:%
%:%296=201%:%
%:%297=202%:%
%:%298=203%:%
%:%299=204%:%
%:%300=205%:%
%:%301=206%:%
%:%302=207%:%
%:%303=208%:%
%:%304=209%:%
%:%305=210%:%
%:%306=211%:%
%:%307=212%:%
%:%308=213%:%
%:%309=214%:%
%:%310=215%:%
%:%311=216%:%
%:%312=217%:%
%:%321=220%:%
%:%333=222%:%
%:%334=223%:%
%:%335=224%:%
%:%336=225%:%
%:%337=226%:%
%:%338=227%:%
%:%339=228%:%
%:%340=229%:%
%:%341=230%:%
%:%342=231%:%
%:%343=232%:%
%:%344=233%:%
%:%345=234%:%
%:%346=235%:%
%:%347=236%:%
%:%348=237%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{whykant}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Why Kantian Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this thesis, I automate Kantian ethics. In 2006, Powers posited that deontological theories are 
attractive candidates for automation because rules are generally computationally tractable \cite[1]{powers}. 
Intuitively, algorithms are rules or procedures for problem solving and Kantian ethics offer one such 
procedure for the problem of making ethical judgements. I will make this intuition precise by
arguing that Kantian ethics is natural to formalize because evaluating a maxim requires little additional
data about the world and a maxim is relatively easy to represent to a computer. All 
ethical traditions have debates that an automated ethical system will need to take a stance on, but these debates
are less frequent and less controversial for Kantian ethics than for consequentialism and 
virtue ethics.

I do not aim to show that Kantian ethics is the only tractable theory to automate and
to present a comprehensive overview of all consequentialist or virtue ethical theories. Instead, I 
present a sample of some approaches in each tradition and argue that Kantian ethics is more straightforward 
to formalize than these approaches. Future work could and should address the challenges I outline in 
this section. The more ethical theories that computational tools can handle, the more valuable 
computational philosophy becomes both for philosophers and for AI agents. Insofar as my project serves 
as an early proof-of-concept for computational ethics, I choose to automate an ethical theory that 
poses fewer challenges than others.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Kantian Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
\textbf{Crash Course on Kantian Ethics}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Kant's theory is centered on practical reason, which is the kind of reason that we 
use to decide what to do. In \emph{The Groundwork of the Metaphysics of Morals}, Kant's most influential 
text on ethics, he explains that rational beings are unique because we can act ``in accordance with 
the representations of laws" \cite[4:412]{groundwork}. In contrast, a ball thrown into the air acts 
according to the laws of physics themselves. It cannot ask itself, ``Should I fall back to the ground?" 
It simply falls. A rational being on the other hand, can ask themselves ``Should I act on this reason?" 
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on" \cite[100]{sources}. 
Rational beings are set apart by this reflective capacity. A rational being's behavior is purposive and 
their actions are guided by practical reason. They have reasons for acting, even when these reasons may be 
opaque to them. This operation of practical reason is what Kant calls the will.

@@

The first formulation of the categorical imperative is the
Formula of Universal Law (FUL), which reads, ``act only according to that maxim through which you can 
at the same time will that it become a universal law" \cite[34]{groundwork}. This formulation
generates the universalizability test, which tests the moral value of a maxim by 
imagining a world in which it becomes a universal law and attempting to will the maxim in that world.
If there is a contradiction in willing the maxim in a world in which everyone universally wills the maxim,
the maxim is prohibited. 
Velleman presents a concise argument for the FUL. He argues that reason is universally shared among reasoners. For 
example, all reasoners have equal access to the arithmetic logic that shows that ``2+2=4" \cite[29]{velleman}. The chain of 
reasoning that makes this statement true is not specific to any person, but is universal across people. 
Therefore, if I have sufficient reason to will a maxim, so does every other rational agent. There is 
nothing special about the operation of my practical reason that other reasoners don't have access to. 
Practical reason is shared, so in adopting a maxim, I implicitly state that all reasoners
across time also have reason to adopt that maxim. Therefore, because I act on reasons, I must obey the 
FUL. Notice that this fulfills the above criterion for a categorical imperative: the FUL is derived from 
a property of practical reason itself and thus derives authority from the will's authority over itself, 
as opposed to some external authority.

The second formulation of the categorical imperative is the formula of humanity (FUH): ``So act that you use humanity, 
in your own person, as well as in the person of any other, always at the same time as an end, never merely 
as a means." \cite[41]{groundwork}. This formulation is often understood as requiring us to 
acknowledge and respect the dignity of every other person. The third formulation of the categorical 
imperative is the formula of autonomy (FOA), which Korsgaard summarizes in her introduction to the Groundwork 
as, ``we should so act that we may think of ourselves as legislating universal laws through our 
maxims" \cite[28]{korsgaardintro}. While closely related to the FUL, the FOA presents morality as the activity of 
perfectly rational agents in an ideal ``kingdom of ends," guided by what Kant calls the ``laws of freedom."

The above is not meant to serve as a full defense of Kant's ethical theory, as that is outside the scope
of this thesis. Instead, I briefly reconstruct an argument for Kant's ethical theory in the hopes 
of offering context for the implementation of the FUL I present later in the thesis. Additionally, understanding 
the structure of Kant's argument also reveals facts about his theory that make it an ideal candidate 
for formalization.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\textbf{Ease of Automation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Kantian ethics is natural to formalize because the categorical imperative, particularly the FUL, 
is a formal principle of practical reason. In other words, the FUL is a property of reason related 
to the form or structure of a maxim. It has nothing to do with the circumstances of behavior 
(beyond those included in a maxim), the agent's mental state, or any other contingent facts. Instead, 
it is purely a property of the proposed principle for action. This formalism makes Kantian ethics an 
attractive candidate for formalization. While other ethical theories often rely on many facts about 
the world or the actor, Kantian ethics simply relies on the form of a given maxim. A computer evaluating 
a maxim doesn't require any knowledge about the world beyond what is contained in a maxim. A maxim 
is the only input that the computer needs to make a moral judgement. Automating 
Kantian ethics merely requires making the notion of a maxim precise and representing it to the computer. 
This distinguishes Kantian ethics from consequentialism and virtue ethics, which, as I will argue below, 
require far more knowledge about the world or the agent to reach a moral decision.

@@

I do not argue that Kantian ethics is the only theory possible to formalize, but instead that Kantian
ethics is relatively easier to formalize. Indeed, like any ethical theory, the Kantian tradition contains 
debates that philosophers disagree on. In this project, I assume stances on debates about
what exactly constitutes the form of a maxim and the correct interpretation of the Formula of 
Universal Law. Those who disagree with my stances will not trust the judgements of my system. Unlike
consequentialism or virtue ethics, these debates are considered close to settled in the Kantian literature 
\cite{ebelsduggan}. Because such debates are fewer and less controversial than those in virtue ethics
or consequentialism, Kantian ethics is relatively easier to formalize.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Consequentialism%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
A consequentialist ethical theory is, broadly speaking, any ethical theory that evaluates an action by evaluating 
its consequences. \footnote{There is long debate about what exactly makes an ethical theory consequentialist \cite{consequentialismsep}. 
For this paper, I will focus on theories that place the moral worth of an act in its the consequences.} For example, 
utilitarianism is a form of consequentialism in which the moral action 
is the action that results in the best consequences or produces the most good \cite{utilsep}. The focus
on the consequences of action distinguishes consequentialists from Kantians, who derive the moral worth
of an action from the maxim that is acted on. Some debates in the consequentialist tradition include 
which consequences of an action matter, what exactly constitutes a ``good" consequence, and how we can 
aggregate the consequences of an action over all the individuals involved.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\textbf{Which Consequences Matter}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Because consequentialism evaluates the state of affairs following an action, this kind of ethical 
reasoning requires more knowledge
about the state of the world than Kantian ethics. Under a naive version of consequentialism, evaluating 
an action requires perfect knowledge of all consequences following an action. This requires that an 
automated ethical system somehow collect all of the infinite consquences following an action, a likely
impossible task. Moreover, compiling this database of consequences requires 
answering difficult questions about which consequences were actually caused by an action.\footnote
{maybe cite the debate about difficulties in determining causation?}

@@

The challenge of understanding and representing the circumstances of action is not unique to consequentialism,
but is particularly acute for consequentialism. Kantian ethicists robustly debate which circumstances 
of an action are ``morally relevant" and should be included in the formulation of the maxim.\footnote{Powers 
\cite{powers} identifies this as a challenge for automating Kantian ethics and briefly sketches 
solutions from O'Neill \citet{constofreason}, Silber \citet{silber}, and Rawls \citet{rawlsconstructivism}. } Those 
using my system will need to automatically or manually formulate maxims, a process that involves 
common-sense reasoning to determine which circumstances, act, and goal tuples the agent is adopting.
\footnote{For more on the challenge of parsing of ethical dilemmas into maxims, see Section AI Ethics}
Because Kantian ethics merely evaluates a single maxim, the surface of this debate is much smaller
than the debate about circumstances and consequences in a consequentialist system. An automated 
consequentialist system must make such judgements about the act itself, the circumstances in which 
it is performed, and the circumstances following the act. All ethical theories relativize
their judgements to the situation in which an act is performed to some extent, but consequentialism
requires far more knowledge about the world than Kantian ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\textbf{Theory of the Good}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Another debate that an automated consequentialist reasoner would need to take a stance on is
the question of what qualifies as a ``good consequence," or what the theory of the good is. Hedonists associate
good with the presence of pleasure and the absence of pain. Preference utiliarians believe that good is 
the satisfaction of desires and is thus derived from individuals' preferences, as opposed to some
sensation of pleasure or pain. Other consequentialists adopt a pluralistic theory of value, under which 
many different kinds of things are good for different reasons. For example, Moore values beauty and truth 
and other pluralists value justice, love, and freedom \cite{moorepe}. Welfare utilitarians value a person's 
welfare and utilitarians of right value states of affairs in which respect for some set of
rights is maximized \cite{consequentialismsep}.

Most of the above theories of good require that a moral reasoner understand complex features about
individuals' preferences, desires, or sensations in order to evaluate a moral action, making automated
consequentialist ethics difficult. Regardless of the theory of the good, a consequentialist ethical 
reasoner needs to evaluate a state of affairs, which encompass each involved individual's pleasure, 
preferences, welfare, freedom, rights, or whatever other criteria make a state good. This requires
judgements about whether or not a state of affairs actually satisifes the relevant criteria for goodness. 
These judgements are difficult and debateable, and any consequentialist decision requires many of 
these judgements for each individual involved. As systems become more complex and involve more people and 
more acts, making these judgements quickly becomes difficult, posing a scaling challenge for a 
consequentialist ethical reasoner. Perfect knowledge of tens of thousands of people's pleasure or 
preferences or welfare or rights is impossible. Either a human being 
assigns values to states of affairs, which quickly becomes difficult to scale, or the machine does, 
which requires massive common-sense, increases room for doubting the system's judgements, and simplifies
the judgements. This is a tractable problem, but is much more difficult than the Kantian task of formulating
and evaluating a maxim.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\textbf{Aggregation}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Once an automated consequentialist agent assigns a goodness measurement to each person in a state of affairs, it 
must also calculate an overall goodness measurement for the state of affairs. One approach to assigning
this value is to aggregate each person's individual goodness score into one complete score for a state. 
For example, under a simple welfare model, each person is assigned a welfare score and the total 
score for a state of affairs is the sum of the welfare scores for each involved person.
The more complex the theory of the good, the more difficult this aggregation becomes. For example, 
pluralistic theories struggle to explain how different kinds of value can be compared \cite{consequentialismsep}. 
How do we compare one unit of beauty to one unit of pleasure? Subjective theories of the good, such 
as those focused on the sensation of pleasure or an individual's preferences, present difficulties in 
comparing different people's subjective measures. Resolving this debate requires that the automated reasoner 
choose one specific aggregation algorithm, but those who disagree with this choice will not trust 
the reasoner's moral judgements. Moreover, for complex theories of the good, this aggregation algorithm
may be complex and may require a lot of data.

To solve this problem, some consequentialists reject aggregation entirely and instead prefer wholistic
evaluations of a state of affairs. While this approach no longer requires that a reasoner define an 
aggregation algorithm, the reasoner still needs to calculate a goodness measurement for a state of 
affairs. Whereas before the reasoner could restrict analysis to a single person, the algorithm must now 
evaluate an entire state wholistically. Evaluating the goodness of an entire state of affairs is more complicated
than evaluating the goodness of a single person. As consequentialists modulate between aggregation 
and wholistic evaluation, they face a tradeoff between the difficulty of aggregation and the complexity 
of goodness measurements for large states of affairs. This tradeoff also holds for an automated
consequentialist moral agent. Such an agent either needs to define an aggregation function, thus opening 
the door to those who disagree with this definition, or needs to evaluate the goodness of entire states
of affairs, which is a complex and data-intensive philosophical and technical challenge.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\textbf{Prior Attempts to Formalize Consequentialism}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
None of the challenges described above are intractable or capture the full literature of 
all variations of consequentialism. I do not argue that consequentialism is 
impossible to automate. Instead, each of the challenges above requires that the developer 
``plant certain flags" and take a stance on certain philosophical debates. Such debates are present in 
any ethical theory, but consequentialism has more such points of difficulty than Kantian ethics and 
is thus more difficult to automate.

@@

\isamarkupsubsubsection{Virtue Ethics%

@@

\textbf{What Is Virtue}%

@@

\textbf{Evaluating Moral Character}%

@@

\textbf{Machine Learning and Virtue Ethics}%

@@

%:%74=31%:%

@@

%:%158=111%:%
%:%162=113%:%
%:%163=114%:%
%:%164=115%:%
%:%165=116%:%
%:%166=117%:%
%:%167=118%:%
%:%168=119%:%
%:%169=120%:%
%:%170=121%:%
%:%171=122%:%
%:%172=123%:%
%:%173=124%:%
%:%174=125%:%
%:%175=126%:%
%:%176=127%:%
%:%177=128%:%
%:%178=129%:%
%:%179=130%:%
%:%180=131%:%
%:%181=132%:%
%:%182=133%:%
%:%183=134%:%
%:%184=135%:%
%:%185=136%:%
%:%186=137%:%
%:%187=138%:%
%:%188=139%:%
%:%189=140%:%
%:%190=141%:%
%:%191=142%:%
%:%192=143%:%
%:%193=144%:%
%:%194=145%:%
%:%203=148%:%
%:%215=150%:%
%:%216=151%:%
%:%217=152%:%
%:%218=153%:%
%:%219=154%:%
%:%220=155%:%
%:%221=156%:%
%:%222=157%:%
%:%223=158%:%
%:%227=160%:%
%:%231=162%:%
%:%232=163%:%
%:%233=164%:%
%:%234=165%:%
%:%235=166%:%
%:%236=167%:%
%:%237=168%:%
%:%238=169%:%
%:%239=170%:%
%:%240=171%:%
%:%241=172%:%
%:%242=173%:%
%:%243=174%:%
%:%244=175%:%
%:%245=176%:%
%:%246=177%:%
%:%247=178%:%
%:%248=179%:%
%:%249=180%:%
%:%250=181%:%
%:%251=182%:%
%:%252=183%:%
%:%253=184%:%
%:%254=185%:%
%:%255=186%:%
%:%256=187%:%
%:%257=188%:%
%:%258=189%:%
%:%259=190%:%
%:%260=191%:%
%:%261=192%:%
%:%262=193%:%
%:%263=194%:%
%:%264=195%:%
%:%265=196%:%
%:%266=197%:%
%:%267=198%:%
%:%268=199%:%
%:%269=200%:%
%:%270=201%:%
%:%271=202%:%
%:%272=203%:%
%:%273=204%:%
%:%274=205%:%
%:%275=206%:%
%:%276=207%:%
%:%277=208%:%
%:%278=209%:%
%:%279=210%:%
%:%280=211%:%
%:%281=212%:%
%:%282=213%:%
%:%283=214%:%
%:%284=215%:%
%:%285=216%:%
%:%286=217%:%
%:%290=220%:%
%:%294=222%:%
%:%295=223%:%
%:%296=224%:%
%:%297=225%:%
%:%298=226%:%
%:%299=227%:%
%:%300=228%:%
%:%301=229%:%
%:%302=230%:%
%:%303=231%:%
%:%304=232%:%
%:%305=233%:%
%:%306=234%:%
%:%307=235%:%
%:%308=236%:%
%:%309=237%:%
